### 2014-01-20 arfi_methods.tex
draft edits and spelling

\subsection{ARFI Imaging}
Experimental ARFI imaging data were acquired using a modified Siemens Acuson
SC2000\texttrademark~ultrasound scanner (Siemens Healthcare, Ultrasound Business Unit,
Mountain View, CA, USA) and the longitudinal array of an Acuson ER7B
transducer.  The ARFI imaging sequence was comprised of
standard B-mode ultrasonic imaging, or tracking beams, and pushing beams. For
each lateral location, two pre-push reference images were acquired, then three
300 cycle pushing pulses were transmitted in rapid succession, focused at 30
mm, 22.5 mm, and 15 mm, respectively, and finally the response of the tissue
was tracked for up to 6ms at a PRF of 8kHz. This pushing strategy is similar to
what has been published by Bercoff \etal~\cite{Bercoff2004}. The 30 mm and 22.5
mm foci pushing pulses were transmitted at 4.6 MHz with a F/2 geometry and the
15 mm focus pushing pulse was transmitted at 5.4 MHz with a F/2.35 geometry to
maintain the same beamwidth (0.67 mm) throughout the region of excitation. A
total of 82 lateral locations were interrogated to cover the 55 mm field of
view, translating 0.67 mm laterally per location.

@@

3D volumetric imaging data was acquired using mechanical rotation of the ER7B
transducer between sequential imaging frames, sweeping across the lateral
extent of the prostate.  This rotation setup utilized a CIVCO
Micro-Touch\texttrademark stabalizer (CIVCO Medical Solutions, Kalona, IA USA)
that allows for complete 6-axis degrees of freedom for manual positioning of
the transducer to sweep through the entire prostate during imaging.  Sequential
imaging frames were $\sim$ 1-2\degree apart, depending on the absolute size of the
prostate.  A custom optical angular feedback transduction circuit utilizing a
reflective linear strip with 500 lines-per-inch (LPI) resolution (US Digital,
Vancouver, WA, USA) was coupled with a 141 oz-in torque stepper motor with
planetary gearbox (Model \#11YPG202S-LW4-R27, Anaheim Automation, Anaheim, CA,
USA) (Figure~\ref{fig:setup_annotated}) to achieve accurate spatial
localization of the imaging frames in the 3D dataset.

@@

>>>>

\subsection{ARFI Imaging}
Experimental ARFI imaging data were acquired using a modified Siemens Acuson
SC2000\texttrademark~ultrasound scanner (Siemens Healthcare, Ultrasound Business Unit,
Mountain View, CA, USA) and the longitudinal array of an Acuson ER7B
transducer.  The ARFI imaging sequence was comprised of
standard B-mode ultrasonic imaging, or tracking beams, and pushing beams. For
each lateral location, two pre-push reference images were acquired, then three
300 cycle pushing pulses were transmitted in rapid succession, focused at 30
mm, 22.5 mm, and 15 mm, respectively, and finally the response of the tissue
was tracked for up to 6 ms at a PRF of 8kHz. This pushing strategy is similar to
what has been published by Bercoff \etal~\cite{Bercoff2004}. The 30 mm and 22.5
mm foci pushing pulses were transmitted at 4.6 MHz with a F/2 geometry and the
15 mm focus pushing pulse was transmitted at 5.4 MHz with a F/2.35 geometry to
maintain the same beamwidth (0.67 mm) throughout the region of excitation. A
total of 82 lateral locations were interrogated to cover the 55 mm field of
view, translating 0.67 mm laterally per location.

@@

3D volumetric imaging data was acquired using mechanical rotation of the ER7B
transducer between sequential imaging frames, sweeping across the lateral
extent of the prostate.  This rotation setup utilized a CIVCO
Micro-Touch\texttrademark stabilizer (CIVCO Medical Solutions, Kalona, IA USA)
that allows for complete 6-axis degrees of freedom for manual positioning of
the transducer to sweep through the entire prostate during imaging.  Sequential
imaging frames were $\sim$1\degree apart, depending on the absolute size of the
prostate.  A custom optical angular feedback transduction circuit utilizing a
reflective linear strip with 500 lines-per-inch (LPI) resolution (US Digital,
Vancouver, WA, USA) was coupled with a 141 oz-in torque stepper motor with
planetary gearbox (Model \#11YPG202S-LW4-R27, Anaheim Automation, Anaheim, CA,
USA) (Figure~\ref{fig:setup_annotated}) to achieve accurate spatial
localization of the imaging frames in the 3D dataset.

@@

The normalized displacement data were scan-converted using a 3D Delaunay
triangulation interpolation method in Matlab (R2012b, The MathWorks, Inc,
Natick, MA, USA).  The scan-converted ARFI imaging data had a final isotropic
voxel size of 0.15 $\times$ 0.15 $\times$ 0.15 mm$^3$.

### 2014-01-20 disc_arfi_mr.tex
updated text formatting
draft edits and spelling

\subsection{ARFI and MR T2WI Agreement}
This study demonstrated that ARFI imaging delineates zonal anatomy of the
prostate well when compared to MR T2WI with only slight overestimation of 
ARFI total and central gland volume (Figure~\ref{fig:mr_arfi_volumes}). 
 Establishing accurate delineation of the prostate's zonal anatomy is 
important because PCa has strong correlations with location in the prostate, 
specifically in the peripheral zone, and accurate capsule delineation will f
acilitate future efforts to register 3D imaging datasets between ARFI and MR imaging.

We did not expect perfect agreement between MR and ARFI imaging volume or tri-axial
 dimensions since each modality applies a different amount of external compression to 
the rectal wall that deforms the prostate different amounts. Pressure on the prostate 
during ARFI imaging is modulated by the urologist when he locks the transducer rotation 
apparatus in place prior to imaging, and deformation of the prostate can be easily visualized
 during this alignment process.  It was our goal to apply as little pressure while still maintaining
 uniform contact throughout the image acquisition. This is critical since poor acoustic coupling to 
the prostate through the rectal wall can greatly diminish ARFI image displacement SNR.

@@

ARFI imaging consistently underestimated the extent of the prostate capsule and 
central gland apex-to-base dimension relative to MR T2WI ((Figure~\ref{fig:mr_arfi_path_axes}(c), 
Table~\ref{tab:mr_arfi_axes_error}).  In fact, it is likely the degree of ARFI estimation relative 
to the actual apex-to-base prostate length is even greater than suggested by the MR comparison as this 
was the dimension with the worst spatial resolution in MR imaging, with a
typical MR T2WI slice thickness of $\sim$3 mm for these studies.  While
higher-resolution MR images could be obtained and analyzed for more precise
comparisons in this imaging dimension, such data was not available for this
study.

>>>>

\subsection{ARFI and MR T2WI Agreement}
This study demonstrated that ARFI imaging delineates zonal anatomy of the
prostate well when compared to MR T2WI with only slight overestimation of ARFI
total and central gland volume (Figure~\ref{fig:mr_arfi_volumes}).
Establishing accurate delineation of the prostate's zonal anatomy is important
because PCa has strong correlations with location in the prostate, specifically
in the peripheral zone, and accurate capsule delineation will facilitate
future efforts to register 3D imaging datasets between ARFI and MR imaging.

We did not expect perfect agreement between MR and ARFI imaging volume or
tri-axial dimensions since each modality applies a different amount of external
compression to the rectal wall, which then deforms the prostate.  Pressure on
the prostate during ARFI imaging is modulated by the urologist when he locks
the transducer rotation apparatus in place prior to imaging, and deformation of
the prostate can be easily visualized during this alignment process.  It was
our goal to apply as little pressure while still maintaining uniform contact
throughout the image acquisition. This is critical since poor acoustic coupling
to the prostate through the rectal wall can greatly diminish ARFI image
displacement SNR.

@@

ARFI imaging consistently underestimated the extent of the prostate capsule and
central gland apex-to-base dimension relative to MR T2WI
((Figure~\ref{fig:mr_arfi_path_axes}(c), Table~\ref{tab:mr_arfi_axes_error}).
In fact, it is likely the degree of ARFI estimation relative to the actual
apex-to-base prostate length is even greater than suggested by the MR
comparison as this was the dimension with the worst spatial resolution in MR
imaging, with a typical MR T2WI slice thickness of $\sim$3 mm for these
studies.  While higher-resolution MR images could be obtained and analyzed for
more precise comparisons in this imaging dimension, such data were not available
for this study.

### 2014-01-20 fig_arfi_segs.tex
draft edits and spelling

\begin{figure}[htb!]
\centering
\includegraphics[width=0.3\textwidth]{zach/Cap_Modeling_Images/3D_Capsule.png} \\
Prostate Capsule \\
\begin{tabular}{ccc}
\includegraphics[width=0.3\textwidth]{zach/Cap_Modeling_Images/Axial_Cap.png} &
\includegraphics[width=0.3\textwidth]{zach/Cap_Modeling_Images/Coronal_Cap.png} &
\includegraphics[width=0.3\textwidth]{zach/Cap_Modeling_Images/Sagittal_Cap.png} \\
Axial B-mode Outline & Coronal B-mode Outline & Sagital B-mode Outline \\
\end{tabular}
\includegraphics[width=0.3\textwidth]{zach/CG_Modeling_Images/3D_CG.png} \\
Prostate Central Gland \\
\begin{tabular}{ccc}
\includegraphics[width=0.3\textwidth]{zach/CG_Modeling_Images/Axial_CG.png} &
\includegraphics[width=0.3\textwidth]{zach/CG_Modeling_Images/Coronal_CG.png} &
\includegraphics[width=0.3\textwidth]{zach/CG_Modeling_Images/Sagittal_CG.png} \\
Axial ARFI Outline & Coronal ARFI Outline & Sagital ARFI Outline \\
\end{tabular}
\caption{Representative ultrasound prostate models from this study.  TOP ROWS:
    Prostate capsule 3D model (magenta) rendered from manual segmentation of
    the B-mode images, with the 3D model outlines superimposed on the central
    axial, coronal and sagital B-mode images.  The primary segmentation plane
    was sagittal; however the axial and coronal planes were used to guide the
    segmentation to insure 3D continuity of the capsule structure.  BOTTOM
    ROWS: Prostate central gland 3D model (blue) rendered from manual
    segmentation of the ARFI images, with the 3D model outlines superimposed on
    the central axial, coronal and sagital ARFI images.  The coronal ARFI
    imaging plane was the primary orientation used for image segmentation, but
    like the capsule, the orthogonal planes were used to insure continuity of
    the central gland in all three dimensions.}
\label{fig:arfi_segs} 
\end{figure}

>>>>

\begin{figure}[htb!]
\centering
\includegraphics[width=0.3\textwidth]{zach/Cap_Modeling_Images/3D_Capsule.png} \\
Prostate Capsule \\
\begin{tabular}{ccc}
\includegraphics[width=0.3\textwidth]{zach/Cap_Modeling_Images/Axial_Cap.png} &
\includegraphics[width=0.3\textwidth]{zach/Cap_Modeling_Images/Coronal_Cap.png} &
\includegraphics[width=0.3\textwidth]{zach/Cap_Modeling_Images/Sagittal_Cap.png} \\
Axial B-mode Outline & Coronal B-mode Outline & Sagittal B-mode Outline \\
\end{tabular}
\includegraphics[width=0.3\textwidth]{zach/CG_Modeling_Images/3D_CG.png} \\
Prostate Central Gland \\
\begin{tabular}{ccc}
\includegraphics[width=0.3\textwidth]{zach/CG_Modeling_Images/Axial_CG.png} &
\includegraphics[width=0.3\textwidth]{zach/CG_Modeling_Images/Coronal_CG.png} &
\includegraphics[width=0.3\textwidth]{zach/CG_Modeling_Images/Sagittal_CG.png} \\
Axial ARFI Outline & Coronal ARFI Outline & Sagittal ARFI Outline \\
\end{tabular}
\caption{Representative ultrasound prostate models from this study.  TOP ROWS:
    Prostate capsule 3D model (magenta) rendered from manual segmentation of
    the B-mode images, with the 3D model outlines superimposed on the central
    axial, coronal and sagittal B-mode images.  The primary segmentation plane
    was sagittal; however the axial and coronal planes were used to guide the
    segmentation to insure 3D continuity of the capsule structure.  BOTTOM
    ROWS: Prostate central gland 3D model (blue) rendered from manual
    segmentation of the ARFI images, with the 3D model outlines superimposed on
    the central axial, coronal and sagittal ARFI images.  The coronal ARFI
    imaging plane was the primary orientation used for image segmentation, but
    like the capsule, the orthogonal planes were used to insure continuity of
    the central gland in all three dimensions.}
\label{fig:arfi_segs} 
\end{figure}

### 2014-01-20 fig_mr_anatomy.tex
draft edits and spelling

\begin{figure}[htb!]
\centering
\begin{tabular}{ccc}
\includegraphics[width=0.29\linewidth]{figs/mr_anatomy/T2_apex} &
\includegraphics[width=0.275\linewidth]{figs/mr_anatomy/T2_midgland} &
\includegraphics[width=0.3\linewidth]{figs/mr_anatomy/T2_base} \\
(a) MR T2WI Apex & (b) MR T2WI Mid-gland & (c) MR T2WI Base \\
\end{tabular}
\caption{Axial T2-weighted MR images of the prostate show the apex (a),
    mid-gland (b), and base (c).  The peripheral zone (PZ) is of higher signal
    intensity than the central gland (CG), the latter which is composed of the
    central zone and the transitional zone. The apex (A) is composed mostly of
    PZ glandular tissue and the urethra is seen at the level of the midgland as
    an inverted ``U'' (b). Note the area of hypointense signal in the
    peripheral zone at the midgland and base (asterisk, b and c), which
    represents a prostatic tumor.  The posterior (P) aspect of the prostate is
    adjacent to the endorectal coil, and the right (R)-to-left (L) extent of
    the prostate is refered to as the lateral-to-lateral axis in the subsequent
    analysis.}
\label{fig:mr_anatomy} 
\end{figure}

>>>>

\begin{figure}[htb!]
\centering
\begin{tabular}{ccc}
\includegraphics[width=0.29\linewidth]{figs/mr_anatomy/T2_apex} &
\includegraphics[width=0.275\linewidth]{figs/mr_anatomy/T2_midgland} &
\includegraphics[width=0.3\linewidth]{figs/mr_anatomy/T2_base} \\
(a) MR T2WI Apex & (b) MR T2WI Mid-gland & (c) MR T2WI Base \\
\end{tabular}
\caption{Axial T2-weighted MR images of the prostate show the apex (a),
    mid-gland (b), and base (c).  The peripheral zone (PZ) is of higher signal
    intensity than the central gland (CG), the latter which is composed of the
    central zone and the transitional zone. The apex (A) is composed mostly of
    PZ glandular tissue and the urethra is seen at the level of the mid-gland as
    an inverted ``U'' (b). Note the area of hypointense signal in the
    peripheral zone at the mid-gland and base (asterisk, b and c), which
    represents a prostatic tumor.  The posterior (P) aspect of the prostate is
    adjacent to the endorectal coil, and the right (R)-to-left (L) extent of
    the prostate is referred to as the lateral-to-lateral axis in the subsequent
    analysis.}
\label{fig:mr_anatomy} 
\end{figure}

### 2014-01-20 fig_mr_arfi_weight.tex
draft edits and spelling

\begin{figure}[htb!]
\centering
\begin{tabular}{lll}
\includegraphics[width=0.3\linewidth]{figs/corr_path_vol_weight_vol} &
\includegraphics[width=0.3\linewidth]{figs/corr_weight_vol} &
\includegraphics[width=0.3\linewidth]{figs/corr_pathVol_vol} \\
(a) Path Weight : Path Volume & (b) Image Volume : Prostate Weight & (c) Image Volume : Path Volume \\
\end{tabular}
%\begin{tabular}{ll}
%\includegraphics[width=0.3\linewidth]{figs/corr_weight_vol_no4} &
%\includegraphics[width=0.3\linewidth]{figs/corr_pathVol_vol_no4} \\
%(d) Image Volume : Prostate Weight (-4) & (e) Image Volume : Path Volume (-4) \\
%\end{tabular}
\caption{Tri-axial pathology measurements were used to make an ellipsoidal
    prostate volume approximation based on gross pathology axis measurements,
    which was moderately well-correlated with the excised prostated weights (a,
    R$^2$ = \pathVolWeightRsq).  T2WI MR (blue, X) showed a moderate
    correlation between the reconstructed volumes and prostate weight (R$^2$ =
    \weightMRrsq), while volumes reconstructed from ARFI images (green, O)
    showed weaker correlation (R$^2$ = \weightARFIrsq) (b).  Weaker
    correlations existed between both T2WI MR and ARFI image volumes and
    approximated ellipsoidal prostate pathology volumes (R$^2$ = \pathVolMRrsq
    and \pathVolARFIrsq, respectively) (c).  One study subject had a very large
    prostate ($>$ 80 g) that was not well visualized by both MR and ARFI
    imaging, and it was excluded from all of the linear regression.  This
    outlier was indicated with magenta circles in the plots.}
\label{fig:mr_arfi_weight}
\end{figure}

>>>>

\begin{figure}[htb!]
\centering
\begin{tabular}{lll}
\includegraphics[width=0.3\linewidth]{figs/corr_path_vol_weight_vol} &
\includegraphics[width=0.3\linewidth]{figs/corr_weight_vol} &
\includegraphics[width=0.3\linewidth]{figs/corr_pathVol_vol} \\
(a) Path Weight : Path Volume & (b) Image Volume : Prostate Weight & (c) Image Volume : Path Volume \\
\end{tabular}
%\begin{tabular}{ll}
%\includegraphics[width=0.3\linewidth]{figs/corr_weight_vol_no4} &
%\includegraphics[width=0.3\linewidth]{figs/corr_pathVol_vol_no4} \\
%(d) Image Volume : Prostate Weight (-4) & (e) Image Volume : Path Volume (-4) \\
%\end{tabular}
\caption{Tri-axial pathology measurements were used to make an ellipsoidal
    prostate volume approximation based on gross pathology axis measurements,
    which was moderately well-correlated with the excised prostate weights (a,
    R$^2$ = \pathVolWeightRsq).  T2WI MR (blue, X) showed a moderate
    correlation between the reconstructed volumes and prostate weight (R$^2$ =
    \weightMRrsq), while volumes reconstructed from ARFI images (green, O)
    showed weaker correlation (R$^2$ = \weightARFIrsq) (b).  Weaker
    correlations existed between both T2WI MR and ARFI image volumes and
    approximated ellipsoidal prostate pathology volumes (R$^2$ = \pathVolMRrsq
    and \pathVolARFIrsq, respectively) (c).  There was on pathology specimen
    that had an erroneous prostate volume and weight recorded, which has been
    excluded from the linear regressions (magenta circles).}
\label{fig:mr_arfi_weight}
\end{figure}

### 2014-01-20 fig_setup_annotated.tex
draft edits and spelling

\begin{figure}[htb!]
\centering
\includegraphics[width=0.75\textwidth]{figs/setup_annotated.png}
\caption{The experimental setup used for ARFI imaging in the operating suite,
    including the Siemens SC2000 scanner with the ER7B endorectal transducer
    sitting in a custom cradle of a CIVCO Micro-Touch\texttrademark stabalizer
    with optical feedback controlling a stepper motor to acquire 3D ultrasound
    imaging data \invivo.  Custom-written Python code was run on the scanner to
    trigger B-mode and ARFI imaging acquisition sequences relative to completed
    rotation increments of $\sim$1\degree.}
\label{fig:setup_annotated} 
\end{figure}

>>>>

\begin{figure}[htb!]
\centering
\includegraphics[width=0.75\textwidth]{figs/setup_annotated.png}
\caption{The experimental setup used for ARFI imaging in the operating suite,
    including the Siemens SC2000 scanner with the ER7B endorectal transducer
    sitting in a custom cradle of a CIVCO Micro-Touch\texttrademark stabilizer
    with optical feedback controlling a stepper motor to acquire 3D ultrasound
    imaging data \invivo.  Custom-written Python code was run on the scanner to
    trigger B-mode and ARFI imaging acquisition sequences relative to completed
    rotation increments of $\sim$1\degree.}
\label{fig:setup_annotated} 
\end{figure}

### 2014-01-20 image_proc_methods.tex
draft edits and spelling

\subsection{Image Zonal Anatomy Segmentation and 3D Model Rendering}
\subsubsection{MR Image Segmentation and Modeling}
Axial MR T2WI images were manually segmented using the smooth polygon tool in
ITK-SNAP~\cite{Yushkevich2006}, using unique labels for the PZ, CG and AFS. The
gland was segmented from base to apex.  The base was identified below the
bladder, and subsequent images were segmented until the last slice with visible
prostatic tissue was identified caudally. The CG, urethra, PZ and AFS were
segmented independently according to their well-established anatomical
characteristics on
T2WI.~\cite{Barentsz2012,Jung2012,Poon1985,Hricak2007,Bonekamp2011} The PZ was
identified by its homogeneous high signal intensity on T2WI, which is usually
similar to that of the nearby periprostatic fat. The CG was visualized and
delineated based on its heterogeneous, lower signal intensity and its location
(Figure~\ref{fig:mr_anatomy}). The urethra was included in the CG segmentation.
When visible, the AFS was identified by its low T2 signal intensity and its
location anterior to the CG.

@@

The 3D models were used to render volume estimates of the PZ and CG, with the
sum of PZ and CG representing the total prostate gland volume. AFS volume
estimates in select MR cases were included in the total prostate gland volume
estimates. Orthogonal tri-axial measurements in the lateral-to-lateral,
apex-to-base, and anterior-to-posterior dimensions were made in slicer by
placing an ROI cube in the model's center and expanding the ROI cube to fit the
maximal model dimensions (Figure~\ref{fig:roi_tool}).

@@

\subsubsection{Ultrasound Image Segmentation and Modeling}
Sagittal B-mode and ARFI ultrasound image stacks were segmented in 3D Slicer
using the \verb+draweffect+ tool to model and measure the prostate capsule and
CG volume and tri-axial dimensions.  The sagittal imaging plane was chosen for
segmentation because it has a greater number of easily identifiable anatomical
structures:  seminal vesicles in the left and right posterior base, urethra in
the mid apex, surrounding fibromuscular tissue in the anterior , and rectum in
the posterior  than either the axial or coronal views. Prostate capsule was
segmented in B-mode as this modality offers both high contrast between
prostatic tissue and surrounding periprostatic fat, and better SNR in the
anterior capsule region than ARFI images. ARFI images were used for CG
segmentation.  Prior to segmentation, both the B-mode and ARFI image stacks
were down-sampled along the sagittal axis by a factor of 20 (i.e., from 360
slices per image stack to 18 slices per image stack) to reduce the number of
manually-segmented frames.

>>>>

\subsection{Image Zonal Anatomy Segmentation and 3D Model Rendering}
\subsubsection{MR Image Segmentation and Modeling}
Axial MR T2WI images were manually segmented using the smooth polygon tool in
ITK-SNAP~\cite{Yushkevich2006}, using unique labels for the PZ, CG and AFS. The
gland was segmented from base-to-apex.  The base was identified below the
bladder, and subsequent images were segmented until the last slice with visible
prostatic tissue was identified caudally. The CG, urethra, PZ and AFS were
segmented independently according to their well-established anatomical
characteristics on
T2WI.~\cite{Barentsz2012,Jung2012,Poon1985,Hricak2007,Bonekamp2011} The PZ was
identified by its homogeneous high signal intensity on T2WI, which is usually
similar to that of the nearby peri-prostatic fat. The CG was visualized and
delineated based on its heterogeneous, lower signal intensity and its location
(Figure~\ref{fig:mr_anatomy}). The urethra was included in the CG segmentation.
When visible, the AFS was identified by its low T2 signal intensity and its
location anterior to the CG.

@@

The 3D models were used to render volume estimates of the PZ and CG, with the
sum of PZ and CG representing the total prostate gland volume. AFS volume
estimates in select MR cases were included in the total prostate gland volume
estimates. Orthogonal tri-axial measurements in the lateral-to-lateral,
apex-to-base, and anterior-to-posterior dimensions were made in 3D Slicer by
placing an ROI cube in the model's center and expanding the ROI cube to fit the
maximal model dimensions (Figure~\ref{fig:roi_tool}).

@@

\subsubsection{Ultrasound Image Segmentation and Modeling}
Sagittal B-mode and ARFI ultrasound image stacks were segmented in 3D Slicer
using the \verb+draweffect+ tool to model and measure the prostate capsule and
CG volume and tri-axial dimensions (Figure~\ref{fig:arfi_segs}).  The sagittal
imaging plane was chosen for segmentation because it has a greater number of
easily identifiable anatomical structures:  seminal vesicles in the left and
right posterior base, urethra in the mid apex, surrounding fibromuscular tissue
in the anterior, and rectum in the posterior. Prostate capsule was segmented in
B-mode as this modality offers both high contrast between prostatic tissue and
surrounding peri-prostatic fat, and better SNR in the anterior capsule region
than ARFI images. ARFI images were used for CG segmentation.  Prior to
segmentation, both the B-mode and ARFI image stacks were down-sampled along the
sagittal axis by a factor of 20 (i.e., from 360 slices per image stack to 18
slices per image stack) to reduce the number of manually-segmented frames.

### 2014-01-20 intro.tex
draft edits and spelling

\section{Introduction}\label{sect:intro}
Prostate cancer (PCa) is the most common non-cutaneous malignancy among men in
the United States and the second leading cause of cancer related
death.~\cite{Howlader2011} Screening with prostate specific antigen (PSA) and
digital rectal examination (DRE) has led to earlier PCa detection, but
performance of these measures is not optimal, leading to imprecise risk
assessment.  Additionally, the random nature of transrectal ultrasound
(TRUS)-guided biopsies can miss or underestimate the burden of
cancer.~\cite{Gupta2013} Earlier disease diagnosis leads to challenges in
deciding optimal management strategies for patients presenting with less
disease burden, whereas missed tumors on TRUS-guided biopsies can result in
inappropriate diagnoses.

The use of non-invasive prostate imaging could lead to improved PCa diagnosis,
risk-stratification, and management.  Magnetic resonance imaging (MRI) has been
available for use in the workup of patients with PCa since the early 1980s, but
early studies on its diagnostic accuracy were heterogeneous.  The more
recent ability to include functional parameters in PCa MRI analysis has yielded
promising results.~\cite{Gupta2013,Hricak2007} Among the MRI modalities
currently used in the study of PCa, it is well established that T2-Weighted
Imaging (T2WI) offers the best assessment of prostate anatomy due to its
ability to delineate prostatic margins, distinguish internal structures and
differentiate among the glandular zones.

>>>>

\section{Introduction}\label{sect:intro}
Prostate cancer (PCa) is the most common non-cutaneous malignancy among men in
the United States and the second leading cause of cancer related
death.~\cite{Howlader2011} Screening with prostate specific antigen (PSA) and
digital rectal examination (DRE) has led to earlier PCa detection, but
performance of these measures is not optimal, leading to imprecise risk
assessment.  Additionally, the random nature of transrectal ultrasound
(TRUS)-guided biopsies can miss or underestimate the cancer
burden.~\cite{Gupta2013} Earlier disease diagnosis leads to challenges in
deciding optimal management strategies for patients presenting with less
disease burden, whereas missed tumors on TRUS-guided biopsies can result in
inappropriate diagnoses.

The use of non-invasive prostate imaging could lead to improved PCa diagnosis,
risk-stratification, and management.  Magnetic resonance imaging (MRI) has been
available for use in the workup of patients with PCa since the early 1980s, but
early studies on its diagnostic accuracy were heterogeneous.  The more
recent ability to include functional parameters in PCa MRI analysis has yielded
promising results.~\cite{Gupta2013,Hricak2007} Among the MRI modalities
currently used in the study of PCa, it is well established that T2-Weighted
Imaging (T2WI) offers the best assessment of prostate anatomy based on its
ability to delineate prostatic margins, distinguish internal structures and
differentiate among the glandular zones.

### 2014-01-20 prostate_anatomy.tex
draft edits and spelling

The prostate gland can be divided from superior-to-inferior into the base,
mid-gland and apex. The urethra enters the prostate proximally at the base and
extends to the mid-gland, at which point the ejaculatory ducts open into the
urethra at the verumontanum.~\cite{Jung2012} The urethra then continues past
the apex and travels through the penis. The prostate can be divided into
glandular and non-glandular components.  The glandular components include the
transitional zone (TZ), central zone (CZ) and peripheral zone (PZ). In the
healthy prostate, each zone contains approximately 5\%, 20\% and 70--80\% of
glandular tissue, respectively.~\cite{Bonekamp2011}, but these ratios can
change significantly in the presence of benign prostatic hyperplasia (BPH). The
non-glandular components include the anterior fibromuscular stroma (AFS) and
the urethra.  In some imaging modalities, the central and transition zones
cannot be visualized as discrete entities, and they are collectively identified
as the central gland (CG).  This CG was one of the prostate structures manually
segmented in MR and ARFI images in this study.

>>>>

The prostate gland can be divided from superior-to-inferior into the base,
mid-gland and apex. The urethra enters the prostate proximally at the base and
extends to the mid-gland, at which point the ejaculatory ducts open into the
urethra at the verumontanum.~\cite{Jung2012} The urethra then continues past
the apex and travels through the penis. The prostate can be divided into
glandular and non-glandular components.  The glandular components include the
transitional zone (TZ), central zone (CZ) and peripheral zone (PZ). In the
healthy prostate, each zone contains approximately 5\%, 20\% and 70--80\% of
glandular tissue, respectively~\cite{Bonekamp2011}, but these ratios can
change significantly in the presence of benign prostatic hyperplasia (BPH). The
non-glandular components include the anterior fibromuscular stroma (AFS) and
the urethra.  In some imaging modalities, the central and transition zones
cannot be visualized as discrete entities, and they are collectively identified
as the central gland (CG).  The CG was one of the prostate structures manually
segmented in MR and ARFI images in this study.

### 2014-01-20 results.tex
draft edits and spelling

Measurements of the prostate total and CG dimensions along the three standard
anatomic axes (apex-to-base, lateral-to-lateral, and anterior-to-posterior)
were made (Table~\ref{tab:mr_arfi_axes}), and the correlation between the
imaging axis measurements was analyzed (Figure~\ref{fig:mr_arfi_path_axes}).
ARFI was most correlated to MR in the lateral-to-lateral axis in both the total
and CG (R$^2$ = \totalLatLatRsq~and \centralLatLatRsq, respectively), with mean
overestimates of \ARFImrTotalLatLatMeanPct~$\pm$ \ARFImrTotalLatLatStdPct\%~and
\ARFImrCentralLatLatMeanPct~$\pm$ \ARFImrCentralLatLatStdPct\%, respectively
(Table~\ref{tab:mr_arfi_axes_error}).  ARFI had moderate correlation with the
total prostate gland and CG axes in the anterior-to-posterior dimension (R$^2$
= \totalAntPostRsq~and \centralAntPostRsq), respectively, with differences of
\ARFImrTotalAntPostMeanPct~$\pm$ \ARFImrTotalAntPostStdPct\% and
\ARFImrCentralAntPostMeanPct~$\pm$ \ARFImrCentralAntPostStdPct\%.  ARFI imaging
had weak correlation with MR images in the apex-to-base dimension (R$^2$ =
\totalApexBaseRsq, total gland and R$^2$ = \centralApexBaseRsq, CG),
with underestimates of \ARFImrTotalApexBaseMeanPct~$\pm$
\ARFImrTotalApexBaseStdPct\% and \ARFImrCentralApexBaseMeanPct~$\pm$
\ARFImrCentralApexBaseStdPct\%, respectively.

>>>>

Measurements of the prostate total gland and CG dimensions along the three
standard anatomic axes (apex-to-base, lateral-to-lateral, and
anterior-to-posterior) were made (Table~\ref{tab:mr_arfi_axes}), and the
correlation between the imaging axis measurements was analyzed
(Figure~\ref{fig:mr_arfi_path_axes}).  ARFI was most correlated to MR in the
lateral-to-lateral axis in both the total gland and CG (R$^2$ =
\totalLatLatRsq~and \centralLatLatRsq, respectively), with mean overestimates
of \ARFImrTotalLatLatMeanPct~$\pm$ \ARFImrTotalLatLatStdPct\%~and
\ARFImrCentralLatLatMeanPct~$\pm$ \ARFImrCentralLatLatStdPct\%, respectively
(Table~\ref{tab:mr_arfi_axes_error}).  ARFI had moderate correlation with the
total prostate gland and CG axes in the anterior-to-posterior dimension (R$^2$
= \totalAntPostRsq~and \centralAntPostRsq), respectively, with differences of
\ARFImrTotalAntPostMeanPct~$\pm$ \ARFImrTotalAntPostStdPct\% and
\ARFImrCentralAntPostMeanPct~$\pm$ \ARFImrCentralAntPostStdPct\%.  ARFI imaging
had weak correlation with MR images in the apex-to-base dimension (R$^2$ =
\totalApexBaseRsq, total gland and R$^2$ = \centralApexBaseRsq, CG), with
underestimates of \ARFImrTotalApexBaseMeanPct~$\pm$
\ARFImrTotalApexBaseStdPct\% and \ARFImrCentralApexBaseMeanPct~$\pm$
\ARFImrCentralApexBaseStdPct\%, respectively.

### 2014-01-20 tab_axis_ratio_over_under.tex
draft edits and spelling

\begin{table}
\centering
\caption{Mean axis ratio differences between ARFI imaging, T2WI MR and
    pathology for the total prostate volume and central gland / zone for the
    imaging modalities.  The three axis ratios analyzed were:
    lateral-to-lateral : apex-to-base (LL:AB), anterior-to-posterio :
    apex-to-base (AP:AB), and lateral-to-lateral : anterior-to-posterior
    (LL:AP).}
\begin{tabular}{|l|l|l|l|l|} \hline
{\bf Image Modality} & {\bf Comparative Measure} & {\bf Total / Central} & {\bf Axes} & {\bf Axis Ratio Difference (\%)} \\ \hline
\input{figs/tab_axis_ratio_over_under_data}
\hline
\end{tabular}
\label{tab:axis_ratio_over_under}
\end{table}

>>>>

\begin{table}
\centering
\caption{Mean axis ratio differences between ARFI imaging, T2WI MR and
    pathology for the total prostate volume and central gland / zone for the
    imaging modalities.  The three axis ratios analyzed were:
    lateral-to-lateral : apex-to-base (LL:AB), anterior-to-posterior :
    apex-to-base (AP:AB), and lateral-to-lateral : anterior-to-posterior
    (LL:AP).}
\begin{tabular}{|l|l|l|l|l|} \hline
{\bf Image Modality} & {\bf Comparative Measure} & {\bf Total / Central} & {\bf Axes} & {\bf Axis Ratio Difference (\%)} \\ \hline
\input{figs/tab_axis_ratio_over_under_data}
\hline
\end{tabular}
\label{tab:axis_ratio_over_under}
\end{table}
